---
title: "p8131_hw10_mp3745"
author: "Matthew Parker"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(KMsurv)
library(survival)
library(survminer)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


## Problem 1

Test H0 : h1(t) = h2(t) for all t, H1 : h1(t)  ̸= h2(t) for some t using the log-rank test.
```{r}
# Enter in the data
data = tibble(
  time = c(4, 12, 15, 21, 23, 2, 6, 8, 10, 19),
  cens = c(1, 0, 1, 0, 1, 1, 0, 0, 1, 1),
  group = c(1, 1, 1, 1, 1, 2, 2, 2, 2, 2)
)

# Test hypothesis with log rank test
survdiff(Surv(time, cens) ~ group, data = data)
```

Since p-value < 0.05, we fail to reject the null hypothesis that the hazard rate functions are the same for both groups at all time t.


## Problem 2

Read in the data
```{r}
data("kidtran")
```

Compare the survival curves for different races in each sex group
```{r}
ggsurvplot(survfit(Surv(time, delta) ~ gender + race, data = kidtran), conf.int = TRUE)
```

Based on the above survival curves, it looks like gender = 2 and race = 2 have the worse survival. The survival curves for the other race/gender groups are fairly similar and better than gender = 2, race = 2.


## Problem 3

Read in the data
```{r}
data("larynx")

# convert stage 2 to single binary column
larynx_bin = larynx %>% 
  mutate(
    stage_1 = as.factor(if_else(stage == 1, 1, 0)),
    stage_2 = as.factor(if_else(stage == 2, 1, 0)),
    stage_3 = as.factor(if_else(stage == 3, 1, 0)),
    stage_4 = as.factor(if_else(stage == 4, 1, 0)),
    stage = as.factor(stage)
  )
```

Fit a proportional hazards model using the covariates Z1, Z2, Z3, Z4 and the interaction term Z1 * Z4
```{r}
ph_mod = coxph(Surv(time, delta) ~ stage_2 + stage_3 + stage_4 + age + stage_2 * age, data = larynx_bin, ties = 'breslow')

summary(ph_mod)
```

From the model, age is not significantly related to time to death, holding stage constant. The hazard ratio between stage 2 and stage 1, while holding age constant is 0.0006. The hazard ratio between stage 3 and stage 1, while holding age constant is 1.862. The hazard ratio between stage 4 and stage 1, while holding age constant is 5.774. There is a significant stage 2 by age interaction term in time to death.

The relative risk of dying for a stage II patient of age 50 as compared to a stage I patient of the same age is `r exp(ph_mod$coefficients[1] + ph_mod$coefficients[5] * 50)`.


